---
title: "R_tutorial"
author: "Ramsey Cervantes"
date: "3/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

The objective of this code is to learn how to import into R an excel file containing marginal estimates generated in Stata. Then reformatting the data in order to have it become organized for plotting using ggplot. Once this has been accomplished, it will then be put into an rmd file and knitted to an html file.

## Reading in Data

```{r reading data}
# read csv file generated from Stata margins
data <- read.csv("Table1_margins.csv")
# remove equal signs from cells
data <- lapply(data, function(x) gsub("=","",x))
# read in data
data <- data.frame(data)
```

## Formatting the Data

```{r format data}
# store column names
n <-c('b','se','t','pvalue','ll','ul','df','crit','eform',rep(c('x','b','se','t','pvalue','ll','ul','df','crit','eform'),3))
# transpose data
data <- t(data[,-1])[,c(1:39)]
# assign column and row names
colnames(data) <- n
rownames(data) <- NULL
# subset the columns we need
data <- data[,c(1,4,5,6,11,14,15,16,21,24,25,26,31,34,35,36)]
# separate data into identifiable groups
male_diabetes <- as.data.frame(data[,c(1:4)])
female_diabetes <- as.data.frame(data[,c(5:8)])
male_heartatk <- as.data.frame(data[,c(9:12)])
female_heartatk <- as.data.frame(data[,c(13:16)])
data <- rbind(male_diabetes, female_diabetes, male_heartatk, female_heartatk)
# generate/populate identifier columns
data$race <- rep(c("White","Black","Other"),4)
data$type <- rep(c("Diabetes","Heart Attack"),each=6)
data$sex <- rep(rep(c("Male","Female"),each=3),2)
data[,c(1,2,3,4)] <- sapply(data[,c(1,2,3,4)], as.numeric)
# separate data by health risk type
data_diabetes <- data[1:6,]
data_heartatk <- data[7:12,]
```

## Generating Plots

```{r create plots}
# generate diabetes risk plot
plot_diabetes <- ggplot(data_diabetes, aes(x=race, y=b,colour=sex,group=sex)) +
  geom_pointrange(aes(x=race, y=b, ymin=ll, ymax=ul),position=position_dodge(0.3),size=1.05)+
  labs(y="Marginal Probabilities", x="Race",title="Risk of Diabetes",colour='', shape='')+theme_bw()
# generate heart attack risk plot
plot_heartatk <- ggplot(data_heartatk, aes(x=race, y=b,colour=sex,group=sex)) +
  geom_pointrange(aes(x=race, y=b, ymin=ll, ymax=ul),position=position_dodge(0.3),size=1.05)+
  labs(y="Marginal Probabilities", x="Race",title="Risk of Heart Attack",colour='', shape='')+theme_bw()
```

## Health Risk Plots

```{r display plots, echo=FALSE}
plot(plot_diabetes)
plot(plot_heartatk)
```
